### 2020-02-13 manuscript_overflow.rmd
remove extra visualization code

---
title: "manuscript overflow"
author: "Nick Gauthier"
date: "2/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
sim2 %>%
  group_by(rep, time) %>%
  summarise(test = mean(payoff))%>%
  summarise(test = mean(test)) %>%
  #summarise(famine = sum(payoff == 0) / 150) %>%
  #summarise(famine = mean(famine)) %>%
  ggplot(aes(test)) +
  geom_histogram()
#ggplot(aes(time, famine)) +
#  geom_line() +
#  theme_classic()

sim2 %>%
  group_by(rep, time) %>%
  summarise(test = mean(payoff))%>%
  ggplot(aes(time, test, color = rep)) +
  geom_point(alpha = .5)
```

```{r}
sim1 %>%
  group_by(time) %>%
  summarise(payoff = mean(payoff)) %>%
ggplot(aes(time,payoff)) +
  geom_point(aes(color = time)) +
  theme_classic()
```

```{r}
library(ggridges)
sim1 %>%
  mutate(time_step = 5 * round(time/5),
         famine = payoff == 0) %>%
  ggplot(aes(wheat_prop, as.factor(time_step))) +
  geom_density_ridges(rel_min_height = 0.005) +
  theme_ridges()
```

make this a function to calculate payoffs for all wheat or all barley strategies as a comparison
```{r}
tibble(drought = droughts,
       barley = if_else(droughts == FALSE, barley_normal, barley_dry),
       wheat = if_else(droughts == FALSE, wheat_normal, wheat_dry),
       equal = (barley + wheat) / 2) %>%
  mutate(time = 1:n()) %>%
  gather(crop, yield, barley:equal) %>%
  ggplot(aes(time, yield, group = crop, color = crop)) +
  geom_point()
```

```{r}
tibble(drought = droughts,
       barley = if_else(droughts == FALSE, barley_normal, barley_dry),
       wheat = if_else(droughts == FALSE, wheat_normal, wheat_dry),
       equal = (barley + wheat) / 2) %>%
  mutate(time = 1:n()) %>%
  gather(crop, yield, barley:equal) %>%
  mutate(famine = yield == 0) %>%
  ggplot(aes(time, famine, group = crop, color = crop)) +
  geom_line()
```